Validation of Visual Statistical Inference, Applied to Linear Models

### 2012-05-22 paper_jasa.tex
changed one word
working on vis inf glossary

\begin{figure}[htp]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{test_category.pdf}
       \caption{Lineup plot ($m=20$) for testing $H_0: \beta_k=0$. When the alternative hypothesis is true the plot of the observed data should show a vertical displacement between box plots. Which plot has the biggest vertical displacement between the boxplots?}
       \label{fig:test_category}
\end{figure}
%\end{figurehere}

The protocol has only been informally tested until now. In the work described in this paper, the lineup protocol is compared head to head with the relevant classical test. Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model. This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference. The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test. The thought is though, that the classical setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark. Testing is done using a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups. The results are compared with those of the classical test.

@@

\section{Definitions and Explanations for Visual Statistical Inference} \label{sec:visual_test} 
Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.

@@

\begin{table*}[hbtp]
\caption{Comparison of visual inference with existing inference technique}
\centering 
\begin{tabular}{llll} 
\hline
  & Mathematical Inference &  Visual Inference \\ %[0.5ex] % inserts table %heading 
\hline
  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta \ne 0$& $H_0: \beta=0$ vs $H_1: \beta \ne 0$\\
 & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\

 Null Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
 & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
 Reject $H_0$ if & observed $T$ is extreme & observed plot is identifiable \\
\hline 
\end{tabular}
\label{tbl:compare}
\end{table*}

>>>>

\begin{figure}[htp]
%\begin{figurehere}
   \centering
       \includegraphics[width=0.95\textwidth]{test_category.pdf}
       \caption{Lineup plot ($m=20$) using side-by-side boxplots for testing $H_0: \beta_k=0$. One of these plots is the plot of the actual data, and the remaining are null plots, produced by simulating data from a null model that assumes $H_0$ is true. Which plot is the most different from the others, in the sense that there is the the largest shift or location difference between the boxplots? *** Make the legend text use a subscript}
       \label{fig:test_category}
\end{figure}
%\end{figurehere}

The protocol has only been informally tested until now. In the work described in this paper, the lineup protocol is compared head to head with the equivalent classical test. Specifically, the lineup is examined in the context of a linear model setting, where we are determining the importance of including a variable in the model. This is not the envisioned environment for the use of the lineup -- actually it is likely the worst case scenario for visual inference. The intended use of lineups is where there is no existing test, and unlikely to be ever any numerical test. The thought is though, that the classical setting provides a benchmark for how well the lineup protocol works under controlled conditions, and will provide some assurance that they will work in scenarios where there is no benchmark. Testing is done using a human-subjects experiment using Amazon's Mechanical Turk \citep{turk}, using simulation to provide controlled conditions for assessing lineups. The results are compared with those of the classical test.

@@

\section{Definitions and Explanations for Visual Statistical Inference} \label{sec:visual_test}

Table \ref{tbl:compare} illustrates the lineup protocol in relation to traditional hypothesis testing. Both start from the same place, the same null hypothesis. The classical test statistic is the $t$-statistic, where the parameter estimate is divided by its standard error. In the lineup, the test statistic is a plot of the data, here side-by-side boxplots are used, because the variable of interest is categorical and takes just two values. In traditional hypothesis testing the value of the test statistic is compared with all possible values of the sampling distribution, the distribution of the statistic if the null hypothesis is true. If it is extreme on this scale then the null hypothesis is rejected. In constrast in visual inference, the plot of the data is compared with plots of a finite number of samples drawn from the null distribution. If the data plot is selected as the most different, then this results in rejection of the null hypothesis.

\begin{table*}[hbtp]
\caption{Comparison of visual inference with existing inference technique. *** The plot for the traditional test is wrong, for two sided, p-value is area to the right and left of $|T(y|$.}
\centering 
\begin{tabular}{llll} 
\hline
  & Traditional Inference &  Lineup Protocol \\ %[0.5ex] % inserts table %heading 
\hline
  Hypothesis & $H_0: \beta=0$ vs $H_1: \beta \ne 0$& $H_0: \beta=0$ vs $H_1: \beta \ne 0$\\
 & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\

 Sampling Distribution & $f_{T(y)}(t); $\begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.55}{\includegraphics{stat_mathematical_test.pdf}} \end{center} \end{minipage} & $f_{T(y)}(t); $ \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.19}{\includegraphics{test_category.pdf}} \end{center} \end{minipage} \\
 & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} & \begin{minipage}[h]{2.5cm} \begin{center} \scalebox{0.2}{\includegraphics{down_arrow.pdf}} \end{center} \end{minipage} \\
 Reject $H_0$ if & observed $T$ is extreme & observed plot is identifiable \\
\hline 
\end{tabular}
\label{tbl:compare}
\end{table*}

Visual tests are non-parametric tests of the form $H_0$: the data plot is visually not distinguishable from a null plot vs $H_1$: the data plot is visually distinguishable from a plot generated consistently with $H_0$.

@@
